### 2016-03-04 paper.tex
Text improvements

\documentclass[10pt,twocolumn]{article}
\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage{amssymb,amsmath}
\usepackage[usenames,dvipsnames]{color}
\usepackage[colorlinks,citecolor=RedViolet,urlcolor=blue]{hyperref}
\usepackage{doi}
\usepackage{setspace}
\usepackage[paperwidth=8.5in, paperheight=11in,top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage[compact]{titlesec}
\usepackage{abstract}
\usepackage[position=t,singlelinecheck=off,justification=raggedleft]{subcaption}
\usepackage[authoryear,square]{natbib}
\usepackage{aas_macros}
\pagestyle{fancy}
\fancyhead[R]{Veibell \thepage}
\fancyhead[L]{}

\newcommand{\vinote}[1]{\textcolor{red}{\textbf{#1}}} %If you want \inote{#} instead of just \note #, for inline notes
\def\vnote#1\par{\textcolor{red}{\textbf{#1}}\\} %Make note bold and red, re-add newline
\newcommand{\req}{\ensuremath{\rho_{eq}}}
\newcommand{\inote}[1]{\textcolor{blue}{\textbf{#1}}} %Inline notes in blue
\def\note#1\par{\textcolor{blue}{\textbf{#1}}\\}

\begin{document}
\title{Relationship between solar wind, $D_{st}$, and plasmasphere trough mass density on one-hour time scales}
\author{
Victoir Veibell\footnote{vveibell@gmu.edu}
\and
R.S. Weigel\footnote{rweigel@gmu.edu}}

\setlength{\parskip}{3ex}
\renewcommand{\labelitemi}{$-$}
\titlespacing{\section}{0pc}{0.2pc}{-1pc}
\titlespacing{\subsection}{0pc}{0.1pc}{-1pc}
\titleformat*{\section}{\normalsize\bfseries}
\titleformat*{\subsection}{\small\bfseries}

\twocolumn[
  \begin{@twocolumnfalse}
\maketitle
\hrule
\begin{abstract}
This paper compares various magnetosphere conditions around the onset of geomagnetic events, defined as decreases in $D_{st}$ below a threshold value or $\rho_{eq}$ above a threshold value.

\vnote More detail

%It confirms results from previous papers that a sudden drop in $D_{st}$ correlates with a spike in equatorial mass density under certain conditions, while adding a level of depth and specification to the applicability of their results. It compares data at 1-hour averages to data at 1-day averages to examine if the trend holds at varying time scales and finds that it does, so long as the data is pre-processed to linearly remove a large time scale $F_{10.7}$ dependence. By then looking at an hourly average of parameters from 24 hours before event onset to 48 hours after, short timescale trends can be discerned.

\vspace{2em}
\end{abstract}
 \end{@twocolumnfalse}
  ]

\saythanks

\section{Introduction}

The plasma trough is the transition region between the high density and low temperature plasmasphere and the lower density and higher energy magnetosphere \citep{Mayr1968ModelMagnetosphereTemperature}.  Inside of the plasmapause, the density is primarily due to ionospheric photoionization; outside of the plasmapuase, the density has contributions from the co-rotating plasmasphere and conveting magnetosphere plasmas.

Although there are well--validated models, as a function of $L$, $MLT$, and geomagnetic activity, of the plasmasphere density (\cite{Gallagher1988EmpiricalModelPlasmasphere}, \cite{LemaireEarthsPlasmasphere}, \cite{Takahashi2006}, and \cite{Takahashi2010}) and plasmapause boundaries (\cite{LemaireEarthsPlasmasphere}, \cite{Moldwin2002ModelPlasmapause}, \cite{OBrien2003EmpiricalPlasmapause}), the dependence of the plasma trough on geomagnetic and solar wind conditions has not been extensively studied until the past decade.

\cite{Takahashi2006} estimated magnetospheric mass density using measurements from the CRRES satellite during a 73-day period in 1991 and found that the average ion mass, $M=\rho/n_e$, correlated with geomagnetic activity, with more negative $D_{st}$ values corresponding to slightly higher average ion mass densities.  They also found a relationship between the average ion mass and 1.5- and 3-day averages of $K_p$, with an increasing correlation as $K_p$ was smoothed to longer timescales.  The CRRES measurements used were made in the MLT range of 12:00 and 18:00 and mostly in the plasma trough (between 2-7$R_E$) on an elliptical, nearly equatorial orbit.

\cite{Denton2006} showed a relationship between $D_{st}$ and the field line distribution of plasma density, and specifically along the same field lines as looked at in later papers (6-8$R_E$). They found that the distribution depends on $L$, \note{finish}.

\cite{Takahashi2010} developed a mass density dataset using measurements from the Space Environment Monitor instruments on the Geostationary Operational Environmental Satellites (GOES) satellites from 1980 through 1992, with most measurements in the range $L=6.8\pm0.2 R_E$. The mass density $\rho$ was estimated using the the Alfv\'en wave velocity relationship, $V_A=B/\sqrt{\mu_0\rho}$, \vinote{Connect to harmonic toroidal frequency} a magnetic field model, and a numerical solution to a wave equation with an ionospheric boundary and the assumption of a zero resistance ionosphere.  The equatorial mass density, $\rho_{eq}$, was derived from the estimated mass density using a power law dependence on the geocentric distance to the field line of the observation, $R$, $\rho=\rho_{eq}L/(R/R_E)$.

\cite{Takahashi2010} noted that downward spikes in the $D_{st}$ index coincided with with significant changes in $\rho_{eq}$ at for $L$ near 6.8~$R_E$. For five storms, two had $\rho_{eq}$ spikes after the $D_{st}$ drop, two had $\rho_{eq}$ spikes before the drop, and one showed little change in $\rho_{eq}$.  A key result was that when daily-averaged measurements were considered, an enhancment in $\rho_{eq}$ appeared on the same day as minimum $D_{st}$, indicating the possibility of a solar wind control over $\rho_{eq}$.

The results of \cite{Takahashi2006} and \cite{Takahashi2010} are consistent with other results on changes in mass density in the plasma trough region.  \cite{Yao2008} studied the relationship between $D_{st}$ and ion number density for different ion species in different regions (ring current and plasma sheet) and found a general correlation during each of the four storms considered.

In this work we consider the dependence of mass density estimates in the \cite{Takahashi2010} and attempt to identify the solar wind processes that drive changes in the plasma trough density.  There are several issues that are addressed in detail: (1) the plasma density measurements are sparse and (2) the dependence on the north/south component of the solar wind magnetic field is expected to be nonlinear. For northward IMF and low solar wind driving of the magnetosphere, the plasmasphere can expand past geosynchronous orbit where measurements of density are made resulting in an increase in observed density that are not due to solar wind driving; for southward IMF, events studies have shown an increase in density associated with strong southward IMF and enhanced magnetospheric activity [\cite{Takahashi2010}; \cite{Yao2008}].  Issue (1) is addressed by considering \note{Finish}.  Issue (2) is addressed by considering epochs and linear correlations independent of the direction of the IMF and also by separation by averages of the northrward and southward IMF.

--

The dependence of the plasmapause location on geomagnetic activity was observed as early as \note{Cite}.  \cite{Gallagher1988EmpiricalModelPlasmasphere} used measurements from the Retarding Ion Mass Spectrometer onboard the Dynamics Explorer 1 satellite to model the plasmasphere density with a relationship dependent on $L$ shell, altitude, geomagnetic latitude, and observations grouped by a weighted average of $K_p$.  The developed statistical model showed significant improvements when a weighted average of $K_p$ over three days was used to group the observations used in the model.

\cite{Lotoaniu1999PlasmaMassDensity} used measurements from the Plasma Wave Experiment instrument on the CRRES satellite and compared in-situ measured plasma mass density to that inferred by ground-based ULF wave measurements from the CANOPUS array. By extending the ground footprints to the plasmatrough, they found agreement between the two methods for certain plasmapause conditions.  \vinote{Mention specifics? Worth including paper?}

\cite{Hamilton1988RingCurrentDevelopment} analyzed a strong geomagnetic storm and compared energy measurements of the ring current to $D_{st}$, along with measurements of ring current compositions before, during, and after the storm. They found $D_{st}$ to be correlated with the inner ring current energy density during and after storm onset, an inward movement of energy density, and a rapid $D_{st}$ recovery from the quick loss of heavier ions.

\section{Data Preparation and Overview}

The parameters $\rho_{eq}$ and $F_{10.7}$ used in this work are from the dataset associated with \cite{Denton} and used in \cite{Takahashi2010}, with data available from 1980 through 1991; all other parameters are from \cite{Kondrashov2014ReconstructionOfGaps} from 1972 through 2013, which are on a 1--hour time grid. $\rho_{eq}$ is the inferred equatorial mass density based on the 3rd harmonic torodial frequency of magnetic field measurements \vinote{More explanation?}.  The smallest cadence for $\rho_{eq}$ values is 10 minutes.  To compute an hourly average over the same interval for which the solar wind parameters were averaged, the median of all $\rho_{eq}$ values in a given hourly range was used.  Fill (NaN) values were used for hours in which no measurements were available.  In cases where $\rho_{eq}$ was available from multiple GOES satellites at the same time in the \cite{Denton} dataset, the value from GOES 6 was selected.  In the dataset, measurements from GOES 6 had the longest time span of coverage.  An overview of the data used is shown in Figure~\ref{AllData}.

In this work, events are defined to occur when $D_{st}$ or $\rho_{eq}$ crosses a threshold value, as indicated in by horizontal dashed lines in the respective panels of Figure~\ref{AllData}.  In finding $\rho_{eq}$ events, all fill values were replaced with linearly interpolated values. \vinote{Not including \req\ events?}

Figure~\ref{ccplot} shows the long-term trends of GOES 6 based $log(\rho_{eq})$ and $F_{10.7}$ computed using the median values in 27-day non-overlapping windows using all non-fill hourly values.  A scatter plot of these two lines is shown in the lower panel of Figure~\ref{ccplot}.  The linear correlation for GOES 6 is found to be $0.94$ matching the value found by \cite{Takahashi2010} who used measurements from all satellites in the time interval of 1980 through 1991.  The corresponding 27-day correlations for GOES 2, 5, and 7 are 0.81, 0.78, and 0.87, and the time range of coverage of measurements from these satellites are 1980-01-01/1983-05-16, 1983-01-01/1987-02-28, and 1983-05-27/1991-08-29 respectively.

\section{Epoch Average Results}

Two types of events are considered. The first is a drop in the $D_{st}$ index \vnote{Cite Kyoto} below the threshold of -50~nT. These types of events are considered in order to determine if there is a well-defined $\rho_{eq}$ dependence on geomagnetic activity/storms as indicated by $D_{st}$ index.  It is expected that the plasmasphere will exhibit two responses near these threshold crossings.  First is the movement of the plasmasphere earthward and the second is a possible increase in density due to enhanced earthward magnetospheric convection.  The movement of the plasmapause boundary has been well documented and parameterized, most recently by \cite{Moldwin2002ModelPlasmapause} and \cite{OBrien2003EmpiricalPlasmapause}.

The second is an increase in $\rho_{eq}$ above a threshold of 40~amu/cm$^3$.  It is expected that these increases will typically occur after period of quiet geomagnetic activity that follow a geomagnetic storm.  In this case, the plasma pause boundary may move outward past the location of measurements and the increase will be due to measurements being made inside the plasmapause.  A second possible reason for the increase could be due to the above-mentioned increase in magnetospheric convention.

\subsection{$D_{st}$ Events}

Our first analysis uses data only from the time interval 1989-1991 that corresponds with the interval used in the analysis in \cite{Takahashi2010}. For both types of events, the hour of the threshold crossing defines the zero epoch time, and for each event $4\cdot24$ hours were considered before each event and $8\cdot24$ hours after each event.  To compute the epoch averages on a daily time scale, the median value of all measurements for all events centered on a window of $\pm 12$ hours of the epoch day was computed, and these averaging windows were shifted in increments of $24$~hours. (Similar results are obtained if we first reduce each event time series to have a 1-day cadence by computing medians in 1-day bins and then compute the medians across events.)  Error bars for each parameter were computed using the standard deviation of the values used in computing the median divided by the square root of the number of values used.

The stack plot in the upper panel of Figure \ref{DailyAveragedDstEvents} shows the epoch averages for these events.  The minimum $D_{st}$ median is $-48$~nT.  Consistent with \cite{Takahashi2010}, $D_{st}$ events correspond to elevated $\rho_{eq}$, although the magnitude of increase observed here is $1.5$ amu/cm$^3$ (from 18 to 19.5~amu/cm$^3$) instead of the value of $\sim 10$ amu/cm$^3$ found in \cite{Takahashi2010}.  The vertical green bars in the $\rho_{eq}$ plot show the number of $\rho_{eq}$ values that were used to compute its median and error bars.  Note that for the non-\req\ medians, the number of measurements used in computing the medians is typically equal to the number of events, which is much larger than the number of \req\ values as we use all available measurements instead of restricting to only values where a non--fill $\rho_{eq}$ value existed. \vinote{Alternate: Note that for $D_{st}$ events, we computed median values using all available data, even when \req\ availability was sparse.}

In the lower panel of Figure \ref{DailyAveragedDstEvents} it is shown that all trends seen in the analysis of 1989-1991 also hold for the entire span of measurements from GOES 6, from 1983-1991, including the magnitude of change in $\rho_{eq}$ across event onset.

It was found that $\rho_{eq}$ enhancements do not clearly correspond to a signature in $D_{st}$, but that there is a trend for such enhancements to occur when $B_{z}$ in the IMF is positive.  This is consistent with \cite{Takahashi2010} who note that very large increases in $\rho_{eq}$ often corresponded to movement of the plasmasphere to beyond geosynchronous distances.

This method finds 669 periods of $D_{st}$ crossing beneath the selected threshold between May 1983 and August 1991 with an average duration of 9 hours and a median duration of 3 hours \vinote{Make sure I defined "duration"}. Figure \ref{HourlyAveragedDstEvents} shows the average values of all events over a window of 24 hours before onset and 48 hours after. Figure \ref{HourlyAveragedDstEvents}a shows events selected by looking for event onsets where $D_{st}$ crossed a threshold of $-40$~nT. The dashed red lines indicate plus and minus one standard deviation of values from all events that were used to compute the median. The final plot in the stack shows both $\rho_{eq}$ and a bar plot of how many valid data points went into the $\rho_{eq}$ average. Since $\rho_{eq}$ comes from a sparser dataset, it has less valid points contributing to the averages than the other parameters in the stack.

Figure \ref{HourlyAveragedDstEvents} shows a definite spike in the $Z$ component of the magnetic field, as well as the defined drop in $D_{st}$, but no obvious change in mass density at an hourly timescale. This points to an issue with only looking at long-timescale trends between density and $D_{st}$, and allows for the possibility that other factors are influencing the long term correlation since there's no obvious connection on a short timescale. One possibility is that, as suggested in \cite{Takahashi2010}, $F_{10.7}$ plays a significant role in driving long term density values which biases the long term correlation of density and $D_{st}$.

Two types of events are considered. The first is a drop in $D_{st}$ below the threshold of $-40$~nT.  The second is an increase in $\rho_{eq}$ above $40$~amu/cm$^3$.

Our first analysis uses data only from the time interval 1989-1991.  In this interval, there were 329 $D_{st}$ events and 70 $\rho_{eq}$ events.  For both types of events, the hour of the threshold crossing defines the zero epoch time, and for each event $8\cdot24$ hours were kept after each event and $4\cdot24$ hours before each event.  To compute the epoch averages on a daily time scale, the median value of all measurements for all events centered window of $\pm 12$ hours was computed, and these windows were shifted in increments of $24$~hours. (Similar results are obtained if we first reduce each event time series to have a 1-day cadence by computing medians in 1-day bins and then compute the medians accross events.)  Error bars for each parameter were computed using the standard deviation of values used in computing the median dividied by the square root of the number of values used.

In the upper panel of \ref{DailyAveragedDstEvents}, $D_{st}$ events are shown.  The minimum $D_{st}$ median is $-48$~nT.  Consistent with \cite{Takahashi2010}, $D_{st}$ events corrspond to elevated $\rho_{eq}$, although the magnitude of increase observed here is on the order of $\sim2$~amu/cm$^3$ instead of $\sim10$~amu/cm$^3$ found in \cite{Takahashi2010}.  The vertical green bars in the $\rho_{eq}$ plot show the number of $\rho_{eq}$ values that were used to compute its median and error bars.  Note that for $D_{st}$, the number of measurements used in computing the medians is much larger as we use all available $D_{st}$ measurements instead of restricting to only values where a non-fill $\rho_{eq}$ value existed.

Both of the results in \ref{DailyAveragedDstEvents} indicate that there is not a statictically significant change in $\rho_{eq}$ on the day before and day after a threshold crossing in $D_{st}$.

Because the dynamics of the magnetopshere take place on a time scale of less than one day, we have also considered hourly averages as shown in the top panel of \ref{HourlyAveragedDstEvents}.  A total of 668 events were found in the interval of May 1983 through August 1991 with an average duration of 9 hours and a median duration of 3 hours.  The error bars were computed in the same way as \ref{DailyAveragedDstEvents}.  In this plot, the overall trend is similar to that found in the bottom panel of \ref{HourlyAveragedDstEvents} - $\rho_{eq}$ is slighly elevated at the time of the threhshold crossing of $D_{st}$ with respect to the values one day before and after, but the elevation is small and in general not statistically significant.

The bottom panel of \ref{HourlyAveragedDstEvents} contains the events shown in the top panel under the constraint that $D_{st}$ remained below -50~nT for at least twelve hours after the event onset time.

and the event study of \cite{Yao2008} which showed ...

The final plot in the stack shows both $\rho_{eq}$ and a bar plot of how many valid data points went into the $\rho_{eq}$ average. Since $\rho_{eq}$ comes from a sparser dataset, it has less valid points contributing to the averages than the other parameters in the stack. A subset of longer duration events will be looked at later.

This figure shows a definite spike in the $Z$ component of the magnetic field, as well as the defined drop in $D_{st}$, but no obvious change in mass density at an hourly timescale. This points to an issue with only looking at long-timescale trends between density and $D_{st}$, and allows for the possibility that other factors are influencing the long term correlation since there's no obvious connection on a short timescale. One possibility is that, as suggested in \cite{Takahashi2010}, $F_{10.7}$ plays a significant role in driving long term density values which biases the long term correlation of density and $D_{st}$.

\section{More specific events}
It's hypothesized that progressively picking more specific event criteria will allow for the possibility of more significant results, at the expense of more bias in the selection process and potentially less overall usefulness of the results. That said, the predictability of extreme events is of definite interest, so an attempt has been made to find some reproducible method of prediction. Looking at events that last longer than 12 hours and events with an onset threshhold less than $-80$nT finds both result in a spike in mass density, but also a definite lack of data availability to the point where that spike may be coming from less than five of the total 211 events.

\subsection{$\rho_{eq}$ Events}

Instead of looking for events based on a threshold of $D_{st}$, we instead looked for a certain amount of change in $\rho_{eq}$ as the basis for event onset. Both raw change and percent change were done in case of bias towards high or low $\rho_{eq}$ periods, respectively, but neither showed any correlation to a change in $D_{st}$.

An alterntive type of event can be defined that corresponds to large increases in $\rho\ $.  Figure~\ref{HourlyAveragedRhoEvents} shows the epoch time series for events in which an increase of $\req\ $ above $20$~amu/cm$^2$ defines the start of the event.  Here we see the expected result that such events tend to be associated with positive $B_z^{IMF}$.  To determine if the sign of $B_z^{IMF}$ influences the $\rho\ $ epoch time series, we separated the events by the average of $B_z^{IMF}$ in a time window of $4$~hours centered on the event start time.  The epoch curves for $B_z^{IMF}\geq=0$ (257) events and $B_z^{IMF}\ge=0$ (139) have no statistically signifiant difference.

\section{F10.7 dependence}
In an effort to analyze the dependence of $\rho_{eq}$ on $F_{10.7}$, a few tests were performed. \cite{Takahashi2010} mention a strong correlation between the two. The long term correlation could be a bias for $D_{st}$'s effects, so a linear model was created, recreating mass density purely from $F_{10.7}$ in the form of $\rho_{eq}(t)=a\cdot F_{10.7}(t)$. This re-created $\rho_{eq}$ shows around a 45\% correlation with the actual $\rho_{eq}$, suggesting a strong influence, while doing the same procedure with $D_{st}$ shows only a 20-25\% correlation.

Figure \ref{f107bin} takes the events and bins them by the value of $F_{10.7}$ at event onset, then looks at both $\rho_{eq}$ and $B_z$. Error bars calculated in the same manner as before are added to the highest and lowest bins. A significant difference is seen in $\rho_{eq}$ between events taking place during periods of high $F_{10.7}$ and low $F_{10.7}$, while a much less significant, but still noticeable, difference is seen in $B_z$. Also of interest is how $\rho_{eq}$ only sharply reacts to $D_{st}$ events during periods of high $F_{10.7}$. The positive relationship seen in \cite{Takahashi2010} could be due to the short time period looked at, and that it coincides with periods of higher $F_{10.7}$.

Looking at GOES 2,5,6, and 7 shows that this post-onset $D_{st}$ spike happens in almost every case for periods of high $F_{10.7}$. Creating a linear model directly relating variables (and combinations of variables) to $\rho_{eq}$ also shows that $F_{10.7}$ most often has the highest correlation with $\rho_{eq}$. A sample of these correlations is shown in Table \ref{perltable}. The correlations were made by training a one-input linear model of 5-hour post-onset medians of each variable (or variables) against the median of $\rho_{eq}$ at, and four hours after, event onset. This was done by randomly sampling half of the data, generating the model, and testing against the other half, doing this process 100 times and then taking the median correlations.

\input{"tables/CCperltable.tex"}

\section{Appendix}
\subsection{Bias}
While attempting to reproduce Figure 11 from \cite{Takahashi2010}, an ambiguity in data-handling was found. It is unknown how they got from hourly $\rho_{eq}$ to daily medians, whether in one step or two steps, and whether the hourly medians are an hour ahead of each hour grid point, or a median of points a half hour to either side, so attempts were made to reproduce all possibilities and compare in order to find a data-handling method with the least bias. These attempts are shown in Figure \ref{blockmedian}. This also shows the effect of only considering events with a minimum before noon. All show a similar spike in $\rho_{eq}$ at the time of $D_{st}$ minimum, though none quite reproduce the exact values seen in \cite{Takahashi2010}. Figure \ref{allstorms} shows where the median falls for all of the found events during that period, as a sanity check for why the prior work would get median values nearing $30$~amu/cm$^3$. Both authors of this paper conducted independent trials to confirm that the data and analysis don't quite yield the results of Figure 11 in \cite{Takahashi2010} with the processes described in that paper, but the same general trend is still seen.

To check for potential bias in the data as used by our analysis, we checked how the data availability varied with hour, shown in Figure \ref{nanperhour}. We also looked at how storm conditions affected data availability. Both a two sample t-test for difference in means and a Wilcoxon rank sum test showed significance at the 1\% level that $D_{st}$ during periods with available data was of a different distribution than $D_{st}$ during periods missing data. When testing for other significant $D_{st}$ differences, the Wilcoxon test was significant while the t-test was not in pre-noon vs post-noon distributions, as well as $\rho_{eq}>40$ vs $\rho_{eq}<40$ distributions indicating perhaps an increase in variability without an increase in mean.

While the GOES Satellites tend to keep an average geostationary distance of around $L=6.8R_E$ shown in \cite{Takahashi2010}, the actual plasmapause location varies significantly, as shown in \cite{OBrien2003}. This means that during periods of large $D_{st}$ the plasmapause may be far from the point of measurement of $\rho_{eq}$ creating a discord in the correlation. \cite{Gallagher2000} provides a model for $\rho_{eq}$ at a range of L-shells and a brief discussion of the elemental contributions.

\newpage
\footnotesize
\bibliographystyle{plainnat}
\bibliography{paper}

\clearpage
\section{Figures}

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.45]{paperfigures/alldata-GOES6-1983-1991.eps}
\caption{Overview of data used in this study. The top four panels show parameters from \cite{Kondrashov2014ReconstructionOfGaps} and the bottom panel contains $\rho_{eq}$ based on GOES 6 measurements from \cite{Denton} after interpolation and averaging described in the text. Dashed horizontal lines in the $D_{st}$ and $\rho_{eq}$ panels indicate sample event cutoff thresholds of $D_{st} = -50$~nT and $\rho_{eq} = 40$~amu/cm$^3$}.
\label{AllData}
\end{figure}

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.40]{paperfigures/F107MD27d-GOES6.eps}
\includegraphics[scale=0.40]{paperfigures/ccplot-GOES6.eps}
\caption{Top: 27-day boxcar averages of $F_{10.7}$ and $\rho_{eq}$ from GOES 6. Bottom: Correlation between $\log(\rho_{eq})$ and $F_{10.7}$ using hour, day, and 27-day averaging intervals; compare to \cite{Takahashi2010} Fig.~14.}
\label{ccplot}
\end{figure}
\clearpage

\begin{figure}[tp!]
\centering
\includegraphics[scale=0.40]{paperfigures/stormavs-dst-50-tak.eps}
\rule[1ex]{5cm}{1pt}
\includegraphics[scale=0.40]{paperfigures/stormavs-dst-day.eps}
\caption{$D_{st}$ events from GOES 6 using daily medians. Top: Events in the interval 1989-1991; compare to \cite{Takahashi2010} Fig.~11. Bottom: Events in the interval 1983-1991.}
\label{DailyAveragedDstEvents}
\end{figure}

\begin{figure}[tp!]
\centering
\includegraphics[scale=0.40]{paperfigures/stormavs-dst.eps}
\rule[1ex]{5cm}{1pt}
\includegraphics[scale=0.40]{paperfigures/stormavs-dd12.eps}
\caption{$D_{st}$ events from GOES 6 using hourly medians. Top: Events in the interval 1983-1991; compare to bottom panel {DailyAveragedDstEvents}. Bottom: Same as Top except for constraint that $D_{st}$ stayed below $-50$~nT for at least 12~hours after crossing below $-50$~nT.}
\label{HourlyAveragedDstEvents}
\end{figure}

\clearpage

\begin{figure}[htp!]
	\centering
	\includegraphics[scale=0.40]{paperfigures/stormavs-mass-gt20-GOES6.eps}
\caption{$\req\ > 20$~ amu/cm^3 events from GOES 6 using hourly medians. As described in the text, there is no statistically significant dependence of the $\req\ $ plot on $B_z$ in the time interval of $\pm 2$~hours around the event.}
\label{HourlyAveragedRhoEvents}
\end{figure}

\begin{figure}[tp!]
\centering
\includegraphics[scale=0.40]{paperfigures/HighLowF107rhoeq.eps}
\includegraphics[scale=0.40]{paperfigures/HighLowF107Bz.eps}
\includegraphics[scale=0.40]{paperfigures/HighLowF107valid.eps}
\caption{Binning events by $F_{10.7}$ value at onset for $\rho_{eq}$ and, in the middle panel, $B_z$. Lower panel shows number of valid points in each bin}
\label{f107bin}
\end{figure}

\clearpage

\section{Appendix}
\hfill

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.40]{paperfigures/stormavs-d80.eps}
\rule[1ex]{5cm}{1pt}
\caption{(a) Same as Figure \ref{HourlyAveragedDstEvents}(a) except with constraint that $D_{st}$ crossed below $-80$~nT }
\label{Dspec}
\end{figure}

\clearpage

\begin{figure}[htp!]
\includegraphics[scale=0.45]{paperfigures/HighLowF107rhoeq-Dst50-GOES6-1983-1991.eps}
\includegraphics[scale=0.45]{paperfigures/HighLowF107rhoeq-Dst50-GOES7-1987-1991.eps}
\caption{GOES Satellite 6 and 7}
\label{GOES67}
\end{figure}

\begin{figure}[htp!]
\includegraphics[scale=0.45]{paperfigures/HighLowF107rhoeq-Dst50-GOES2-1981-1983.eps}
\includegraphics[scale=0.45]{paperfigures/HighLowF107rhoeq-Dst50-GOES5-1983-1987.eps}
\caption{GOES Satellite 2 and 5}
\label{GOES25}
\end{figure}
\clearpage

\begin{figure}[htp!]
\includegraphics[scale=0.45]{paperfigures/HighLowF107rhoeq-Dst50-GOES6-1989-1991.eps}
\includegraphics[scale=0.45]{paperfigures/HighLowF107rhoeq-Dst80.eps}
\caption{Top: GOES 6 from 89-91 (compare to Takahashi 2010). Bottom:  $D_{st}<-80$~nT.}
\end{figure}

\begin{figure}[htp!]
\includegraphics[scale=0.45]{paperfigures/HighLowF107rhoeq-Dst50-GOES6-1983-1987.eps}
\includegraphics[scale=0.45]{paperfigures/HighLowF107rhoeq-Dst50-GOES6-1987-1991.eps}
\caption{GOES 6 from 83-87 and 87-91}
\end{figure}
\clearpage

\hfill
\begin{figure}[htp!]
\includegraphics[scale=0.45]{paperfigures/HighLowF107rhoeq-Dst30.eps}
\includegraphics[scale=0.45]{paperfigures/HighLowF107rhoeq-Dst30-valid.eps}
\caption{$D_{st} < -30$ nT events, $\rho_{eq}$ binned by $F_{10.7}$. Lower panel: valid hourly points}
\end{figure}
\hfill
\begin{figure}[htp!]
\includegraphics[scale=0.45]{paperfigures/HighLowF107rhoeq-rhoeq40.eps}
\includegraphics[scale=0.45]{paperfigures/HighLowF107rhoeq-rhoeq40-valid.eps}
\caption{$\rho_{eq} > 40$ amu/cm$^3$ events, $\rho_{eq}$ binned by $F_{10.7}$}
\end{figure}
\clearpage

\begin{figure}[htp!]
\includegraphics[scale=0.40]{paperfigures/HighLowF107rhoeq-AE300.eps}
\includegraphics[scale=0.40]{paperfigures/HighLowF107rhoeq-AE300-valid.eps}
\includegraphics[scale=0.35]{paperfigures/stormavs-AE.eps}
\caption{Top: $AE > 400$ nT events, $\rho_{eq}$ binned by $F_{10.7}$ Bottom: averages of $AE > 400$ events}
\end{figure}
\hfill
\begin{figure}[htp!]
\includegraphics[scale=0.45]{paperfigures/HighLowDstrhoeq-AE400.eps}
\includegraphics[scale=0.45]{paperfigures/HighLowDstrhoeq-Dst50.eps}
\caption{Top: $AE > 400$ nT events. Bottom: $D_{st} < -50$ nT events. Both $\rho_{eq}$ binned by $D_{st}$.}
\end{figure}
\clearpage

\begin{figure}[htp!]
\includegraphics[scale=0.45]{paperfigures/HighLowF107rhoeq-random.eps}
\includegraphics[scale=0.45]{paperfigures/stormavs-random.eps}
\caption{Top: $\rho_{eq}$ binned by $F_{10.7}$ for randomly selected times. Bottom: averages of randomly selected times}
\end{figure}

\begin{figure}[htp!]
\includegraphics[scale=0.45]{paperfigures/HighLowDstBz-Dst50.eps}
\includegraphics[scale=0.45]{paperfigures/HighLowDstBz-AE400.eps}
\caption{Top: $B_z$ binned by $D_{st}$ for $D_{st} < -50$ nT events. Bottom: $B_z$ binned by $D_{st}$ for $AE > 400$ nT events}
\end{figure}
\clearpage

\begin{figure}[htp!]
\includegraphics[scale=0.45]{paperfigures/rhoMLT.eps}
\includegraphics[scale=0.45]{paperfigures/rhoLT.eps}
\caption{Median $\rho_{eq}$ vs magnetic local time and universal time}
\end{figure}

\begin{figure}[htp!]
\includegraphics[scale=0.45]{paperfigures/DstLT.eps}
\includegraphics[scale=0.45]{paperfigures/AELT.eps}
\caption{Median $D_{st}$ vs universal time, and median AE vs universal time}
\end{figure}
\clearpage

\end{document}